Type Classes for Mathematics in Type Theory

### 2010-06-28 jar.tex
Fix awkward style in abstract.

The base of our development consists of type classes representing a standard algebraic hierarchy, as well as portions of category theory and universal algebra. We build upon this foundation a set of mathematically sound abstract interfaces for different kinds of numbers, succinctly expressed using categorical language and universal algebra constructions. Strategic use of type classes lets us support these high-level theory-friendly definitions while still enabling efficient implementations unhindered by gratuitous indirection or conversion or projection.

@@

Fields declared with \lstinline|:>| are registered as hints for instance resolution, so that in any context where \lstinline|(A, =, +, 0, *, 1)| is known to be a \lstinline|SemiRing|, \lstinline|(A, =, +, 0)| and \lstinline|(A, =, *, 1)| are automatically known to be \lstinline|CommutativeMonoid|s (and so on, transitively, because instance resolution is recursive). In our hierarchy, these ``substructures'' by themselves establish the inheritance diagram in Figure~\ref{inheritancediagram}. % todo: for some reason the figure number reference is wrong here..

>>>>

The base of our development consists of type classes representing a standard algebraic hierarchy, as well as portions of category theory and universal algebra. On this foundation we build a set of mathematically sound abstract interfaces for different kinds of numbers, succinctly expressed using categorical language and universal algebra constructions. Strategic use of type classes lets us support these high-level theory-friendly definitions while still enabling efficient implementations unhindered by gratuitous indirection or conversion or projection.

@@

Fields declared with \lstinline|:>| are registered as hints for instance resolution, so that in any context where \lstinline|(A, =, +, 0, *, 1)| is known to be a \lstinline|SemiRing|, \lstinline|(A, =, +, 0)| and \lstinline|(A, =, *, 1)| are automatically known to be \lstinline|CommutativeMonoid|s (and so on, transitively, because instance resolution is recursive). In our hierarchy, these ``substructures'' by themselves establish the inheritance diagram in Figure~\ref{inheritancediagram}.
